---
title: "Analysis of COVID-era Apple Mobility Data"
author: "Keren Oczon"
date: "10/5/2021"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

## Setup

```{r set-parameters}
# set the state to work with
state_to_analyze <- "Hawaii"

#sanitizing the state name for output files
state_no_spaces <- gsub(state_to_analyze,
                        pattern = " ",
                        replacement = "_")

# set the name of the global data to use
global_file_to_process <- "data/raw_data/applemobilitytrends-2021-09-23.csv"

# extract core name for output file naming use
core_name <- tools::file_path_sans_ext(basename(global_file_to_process))

```

```{r load-packages-functions}
# load the packages needed
library("dplyr")
library("readr")
library("ggplot2")
library("lubridate")
library("devtools")

# load functions needed
source("code/functions/subset_mobility_data_to_state.R")
source("code/functions/tally_cities_counties_of_state.R")
source("code/functions/change_subset_state_data_to_long.R")
```

## Background and Objectives

This rmarkdown file will use the `ggplot2` [@ggplot2] and `dplyr` [@dplyr] packages 
to analyze the Apple maps mobility data from the time before the COVID-19 pandemic
through the initial shutdowns and the continuing changes over the course of the year. 

The goal will be to create a subset to a particular state, tally up the number of 
cities and counties in that state with mobility data, and then to make several plot 
to visualize those data subets. 

The state that this document will analyze is `r state_to_analyze`. 

## Data Processing

```{r subset-state}
# subset out to the chosenn state using the variables set abo
state_subset <- subset_mobility_data_to_state(
  input_file_name = global_file_to_process,
  state_to_subset = state_to_analyze)
```

```{r make-subset-long}
# Convert the subsetted state date from wide to long format
long_subset <- subset_data_wide_to_long(
  input_state_file = paste0("output/subsetted_state/",
                            core_name,
                            "_",
                            state_no_spaces,
                            ".csv"))
```

```{r tallied-subset}
# tally up the numbers of cities and counties in the state that have mobility
# data available
tallied_subset <- tally_cities_counties_of_state(
  input_file_name = paste0("output/subsetted_state/",
                            core_name,
                            "_",
                            state_no_spaces,
                            ".csv"))
```

## Figures

```{r plot-tallied}
# create a grouped barplot of the tallied data to include in the knitted
# document, and save that figure to the output/figures directory as a png
tallied_plot <- ggplot(data = tallied_subset,
                       aes(x = geo_type,
                           y = n,
                           fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Numbers of cities and/or counties in",
                     state_to_analyze,
                     "with mobility data"),
       x = "Type of area",
       y = "Number of areas")

ggsave(plot = tallied_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_tallied_plot.png"))

tallied_plot

```

```{r plot-time-series}
# produce a line plot of relative driving mobility across the state with date
# on the x axis and mean relative mobility on the y axis
timeseries_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarize(mean_mobility = mean(relative_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = mean_mobility)) +
  geom_line() +
  labs(title = paste("Statewide mean relative mobility driving levels in",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relative mobility")

ggsave(plot = timeseries_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_timeseries_plot.png"))

timeseries_plot
```

```{r plot-histogram-mobility-and-transportation-type}
# create a histogram of the different transportation types and their
# mobility data, differentiating them.
histogram_plot <- ggplot(long_subset, aes(x = relative_mobility)) +
  geom_histogram(color = "white") +
  facet_grid(transportation_type ~ .) +
  labs(title = paste("Relative mobility of different transportation",
                     "types in",
                     state_to_analyze,
                     "during COVID"),
       x = "Relative mobility")

ggsave(plot = histogram_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_historgram_plot.png"))

histogram_plot
```


## Session Info

```{r session-info}
devtools::session_info()
```


## Sources Cited

